---
title: "Task"
format: html
editor: visual
---


## Running Code

Data Import

```{r}
library(tidyverse)

# Load dataset
coffee_data <- read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/refs/heads/main/data/2024/2024-05-14/coffee_survey.csv")

# Preview structure and data
glimpse(coffee_data)
View(coffee_data)

```

 Coffee Type Categorization: coffe_data

```{r}
library(stringr)
library(dplyr)

# Clean and classify coffee preferences
coffee_data <- coffee_data %>%
  mutate(favorite_cleaned = str_to_lower(str_trim(favorite))) %>%  # új, tisztított oszlop
  mutate(favorite_grouped = case_when(
    favorite_cleaned %in% c("black", "drip", "espresso", "americano", "french press", "ristretto", "regular drip coffee") ~ "Black",
    favorite_cleaned %in% c("latte", "flat white", "cappuccino", "macchiato", "mocha", "breve", 
                            "coffee with milk", "coffee + milk", "latte macchiato", "dirty chai", "cortado") ~ "With Milk",
    favorite_cleaned %in% c("iced coffee", "cold brew", "iced latte", "iced americano", 
                            "frappuccino", "blended iced coffee", "cold foam cold brew") ~ "Iced",
    TRUE ~ "Other"
  )) %>%
  relocate(favorite_grouped, .after = favorite)

View(coffee_data)


```


Age Group Categorization and Visualization:

```{r}
coffee_data <- coffee_data %>%
  mutate(age_group = case_when(
    age < 18 ~ "<18",
    age >= 18 & age < 25 ~ "18–24",
    age >= 25 & age < 35 ~ "25–34",
    age >= 35 & age < 45 ~ "35–44",
    age >= 45 & age < 55 ~ "45–54",
    age >= 55 ~ "55+",
    TRUE ~ NA_character_
  ))


library(ggplot2)
library(scales)

coffee_data %>%
  filter(!is.na(age_group), !is.na(favorite_grouped)) %>%
  ggplot(aes(x = age_group, fill = favorite_grouped)) +
  geom_bar(position = "fill") +
  scale_y_continuous(labels = percent_format()) +
  labs(
    title = "Favorite Coffee Types by Age Group",
    x = "Age Group",
    y = "Proportion (%)",
    fill = "Coffee Type"
  ) +
  theme_minimal()

```

 Coffee Preference by Gender:

```{r}
coffee_data %>%
  filter(!is.na(gender), !is.na(favorite_grouped)) %>%
  ggplot(aes(x = gender, fill = favorite_grouped)) +
  geom_bar(position = "stack") +
  labs(
    title = "Favorite Coffee Types by Gender (Total Counts)",
    x = "Gender",
    y = "Number of People",
    fill = "Coffee Type"
  ) +
  theme_minimal()

```


Model 1 – age_group only

```{r}
library(nnet)

mod_data1 <- coffee_data %>%
  filter(!is.na(favorite_grouped), !is.na(age_group))

model1 <- multinom(favorite_grouped ~ age_group, data = mod_data1)

```

Model 2 – age_group + gender:

```{r}
mod_data2 <- coffee_data %>%
  filter(!is.na(favorite_grouped), !is.na(age_group), !is.na(gender))

model2 <- multinom(favorite_grouped ~ age_group + gender, data = mod_data2)

```

Model Comparison:

```{r}

# Predictions
pred1 <- predict(model1)
pred2 <- predict(model2)

# Actual labels
actual1 <- mod_data1$favorite_grouped
actual2 <- mod_data2$favorite_grouped


# Confusion matrices
cat("Confusion Matrix – Model 1 (age_group only):\n")
print(table(Predicted = pred1, Actual = actual1))

cat("\nConfusion Matrix – Model 2 (age_group + gender):\n")
print(table(Predicted = pred2, Actual = actual2))

# Accuracy
acc1 <- mean(pred1 == actual1)
acc2 <- mean(pred2 == actual2)

# AIC values
aic1 <- AIC(model1)
aic2 <- AIC(model2)

 Final comparison
cat("\n--- Model Comparison ---\n")
cat("Model 1 (age_group only): Accuracy =", round(acc1, 3), "| AIC =", round(aic1, 1), "\n")
cat("Model 2 (age_group + gender): Accuracy =", round(acc2, 3), "| AIC =", round(aic2, 1), "\n")

```

Assumption Check – Multicollinearity (VIF)

```{r}
library(car)
mod_vif <- glm(I(favorite_grouped == "Black") ~ age_group + gender, 
               data = mod_data2, 
               family = binomial)

vif(mod_vif)

```

Summary of Findings
This report explored patterns in coffee preferences using survey data from the 2024 TidyTuesday project. After cleaning and categorizing free-text responses, we analyzed relationships between preferred coffee type, age, and gender.

Key Observations
Most popular coffee types were categorized as:

Black, With Milk, Iced, and Other.

Age and gender both influence coffee preferences, with noticeable variation across groups:

Younger age groups preferred Iced and With Milk coffee.

Males showed a relatively higher preference for Black coffee.

Statistical Models
We compared two multinomial logistic regression models:

Model	Accuracy	AIC
Model 1: age_group only	0.363	10097.2
Model 2: age_group + gender	0.413	8683.9

Model 2 outperformed Model 1 in both accuracy and model fit.

Including gender improved prediction and better distributed predicted classes.

Confusion matrices revealed that Model 1 tended to overpredict "With Milk", while Model 2 made more balanced predictions.

Assumption Checks
Variance Inflation Factors (VIF) were calculated for Model 2 to check multicollinearity.

All VIF values were near 1, indicating no concerning correlation between predictors.

Conclusion
The inclusion of gender significantly improves the predictive ability of the model. Age and gender together provide meaningful insight into coffee preferences. The methodology and visualizations support robust, interpretable results and reflect well-documented modeling decisions.

